Chaotic dispersal of tidal debris

### 2015-07-22 ms.tex
more todos

\todo{apw}{rename frequency-space structure or something, show the frequency plots for the three ensembles}
\subsection{Part III: Short-time frequency evolution}\label{sec:results3}

@@

As a demonstration of the small-scale frequency modulation, we consider again the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}). Figure~\ref{fig:fdiff-zooms} shows zoomed regions of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits---the pixel positions are now shown relative to the three parent orbits. The frequency diffusion times shown in these grids were computed with two consecutive windows of length 40 orbital periods. The resonant structure around the mildly chaotic orbit (white lines in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic. To resolve the short-time behavior of the frequency diffusion (corresponding to motion across or around resonance layers) we compute the frequencies in a series of rolling windows along each orbit. We use a window with a width equal to 40 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies---again, this is the smallest window size we could choose before \superfreq\ begins to occasionally fail to recover the frequencies for the strongly chaotic orbit.  Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time---plotted are the percent deviations of the frequency ratios from the initial value. The difference between the first window (green +) and the last window (red x) is 50 orbital periods. The frequencies of the weakly chaotic orbit (panel labeled $W$) evolve quickly but are bounded to a small volume with fractional size $\approx$0.3\% (presumably by nearby resonant surfaces). This is comparable to the frequency spread in globular cluster tidal debris (0.1--0.5\%). %\todo{apw}{check numbers in previous 2 sentences}

@@

\subsection{Part IV: Proposed indicator of chaotic density evolution}\label{sec:results4}

We repeat the above analysis of the short-time evolution of frequencies for all orbits in the orbit grid used above (Section~\ref{sec:results1}): for each orbit, we integrate for 80 orbital periods, but compute the frequencies now in a rolling window along this integration time. We fix the window width to 40 orbital periods and shift the window by 0.5 orbital periods between each computation of the frequencies. For each orbit, we have the three frequencies (and amplitudes), $k=1,2,3$, computed at times $j$. We compute the fractional amplitude of the frequency variation for a given orbit as
\begin{equation}
	A_\Omega = \max_{a_k} \left(\frac{\max_j \Omega^{(j)}_{k} - \min_j \Omega^{(j)}_{k}}{2 \, \mean{\Omega^{(j)}_{k}}_j}\right) \label{eq:frac-freq-amp}
\end{equation}\todo{apw}{change to RMS?}
where, as in Equation~\ref{eq:fdtime}, the outer maximum is taken with respect to the amplitude of the frequency in the power-spectrum, $a_k$, and the inner maximum and minimum are taken over time, $j$. This is a measure of the magnitude of the short-time frequency evolution for a single orbit. For tidal debris with small spreads in orbital properties, we expect nearby orbits to have similar behavior; for example, the central marker in Figure~\ref{fig:fdiff-zooms} shows the size of a globular-cluster-like system in configuration space. We therefore expect a small ensemble of orbits in frequency space to expand over short times around even weakly chaotic parent orbits and the debris will therefore appear dynamically hotter in real-space.

@@

Figure~\ref{fig:freqvar_map} shows the grid of initial conditions, now colored by the (fractional) amplitude of the short-time frequency variation, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}). When the fractional excursions of the frequencies are of order $10^{-3}$--$10^{-2}$ or larger, globular cluster debris will appear more diffuse due to enhanced density evolution from fast, small-scale chaotic diffusion. Comparing this to the timescales of Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}, it is clear that the small-scale variations of the frequencies of a progenitor orbit over a given window of time provides a better prediction of the resulting morphology of tidal debris.

@@

Our main results and conclusions are summarized as follows:
\begin{enumerate}
	\item ``stream-fanning'' of tidal streams on weakly chaotic orbits \citep[as seen in simulations by][]{pearson15} occurs due to small-scale chaotic evolution of the fundamental frequencies of the debris star orbits;
	\item the Lyapunov time and frequency diffusion time are powerful indicators of chaos, but do not capture the importance of small-scale chaotic diffusion for the density evolution of small ensembles of orbits (tidal debris);
	\item tidal debris becomes diffuse and thus harder to observe on weakly chaotic orbits when the small-scale chaotic diffusion of the fundamental frequencies has a scale comparable to the internal spread in frequencies of the debris;
	\item the details of the enhanced mixing along weakly chaotic orbits depends on the resonant structure of the potential;
	\item the variance of the fundamental frequencies of an orbit over a given time window is a strong predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
\end{enumerate}
\todo{apw}{make it clear: this is not about how much chaos, just timescale discrepancy}

@@

% Figure ??
\clearpage
\begin{figure}[p]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/fdiff_zooms.png}
\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the four orbits highlighted in that figure and used in Figures~\ref{fig:orbits}--\ref{fig:densities}. As with Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time. Around each orbit, we initialize a new grid of $\approx$16,000 orbits to resolve the small-scale resonant structure around each orbit. The marker at the center indicates the scale (tidal radius) of a $M \approx 10^4~\msun$ globular cluster at the given position in the initial condition grid. The near-resonant orbit (A) is coincident with a high order resonance (white line), the non-resonant orbit (B) is between resonances, the weakly chaotic orbit (C) is surrounded by intricate structures and may be in a region of resonance overlap, and the strongly chaotic orbit (D) is embedded in a region of strong stochasticity.}
\label{fig:fdiff-zooms}
\end{center}
\end{figure}

>>>>

\subsection{Part III: Short-time frequency evolution}\label{sec:results3}

@@

\todo{apw}{clean up}
As a demonstration of the small-scale frequency modulation, we consider again the four orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:orbit-info}). The resonant structure around the mildly chaotic orbit (white lines in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic. To resolve the short-time behavior of the frequency diffusion (corresponding to motion across or around resonance layers) we compute the frequencies in a series of rolling windows along each orbit. We use a window with a width equal to 40 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies---again, this is the smallest window size we could choose before \superfreq\ begins to occasionally fail to recover the frequencies for the strongly chaotic orbit. Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time---plotted are the percent deviations of the frequency ratios from the initial value. The difference between the first window (green +) and the last window (red x) is 50 orbital periods. The frequencies of the weakly chaotic orbit (panel labeled $W$) evolve quickly but are bounded to a small volume with fractional size $\approx$0.3\% (presumably by nearby resonant surfaces). This is comparable to the frequency spread in globular cluster tidal debris (0.1--0.5\%). %\todo{apw}{check numbers in previous 2 sentences}

@@

We therefore expect a small ensemble of orbits in frequency space to expand over short times around even weakly chaotic parent orbits and the debris should therefore appear dynamically hotter in real-space.

\todo{apw}{insert here the bit about the frequency distributions and chaotic evolution orthogonal to freq distributions}

@@

@@

Our main results and conclusions are summarized as follows:
\todo{apw}{revisit this...}
\begin{enumerate}
	\item ``stream-fanning'' of tidal streams on weakly chaotic orbits \citep[as seen in simulations by][]{pearson15} occurs due to small-scale chaotic evolution of the fundamental frequencies of the debris star orbits;
	\item the Lyapunov time and frequency diffusion time are powerful indicators of chaos, but do not capture the importance of small-scale chaotic diffusion for the density evolution of small ensembles of orbits (tidal debris);
	\item tidal debris becomes diffuse and thus harder to observe on weakly chaotic orbits when the small-scale chaotic diffusion of the fundamental frequencies has a scale comparable to the internal spread in frequencies of the debris;
	\item the details of the enhanced mixing along weakly chaotic orbits depends on the resonant structure of the potential;
	\item the variance of the fundamental frequencies of an orbit over a given time window may be a better predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
\end{enumerate}
\todo{apw}{make it clear: this is not about how much chaos, just timescale discrepancy}

@@

% Figure ??
%\clearpage
%\begin{figure}[p]
%\begin{center}
%\includegraphics[width=0.9\textwidth]{figures/fdiff_zooms.png}
%\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the four orbits highlighted in that figure and used in Figures~\ref{fig:orbits}--\ref{fig:densities}. As with Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time. Around each orbit, we initialize a new grid of $\approx$16,000 orbits to resolve the small-scale resonant structure around each orbit. The marker at the center indicates the scale (tidal radius) of a $M \approx 10^4~\msun$ globular cluster at the given position in the initial condition grid. The near-resonant orbit (A) is coincident with a high order resonance (white line), the non-resonant orbit (B) is between resonances, the weakly chaotic orbit (C) is surrounded by intricate structures and may be in a region of resonance overlap, and the strongly chaotic orbit (D) is embedded in a region of strong stochasticity.}
%\label{fig:fdiff-zooms}
%\end{center}
%\end{figure}
